---
title: 'PALMA - Assignment #4'
output: 
  html_document:
    toc: true
    toc_float: true
    df_print: paged
    code_download: true
---

```{r setup, include=FALSE}
#knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning=FALSE)
```

```{r libraries}
library(tidyverse)         # for graphing and data cleaning
library(dbplyr)            # for SQL query "cheating" - part of tidyverse but needs to be loaded separately
library(mdsr)              # for accessing some databases - goes with Modern Data Science with R textbook
library(RMySQL)            # for accessing MySQL databases
library(RSQLite)           # for accessing SQLite databases
library(dplyr)
theme_set(theme_minimal()) # Lisa's favorite theme
```

## Put it on GitHub! - DONE        

Github Link:  https://github.com/apalma127/assignment-4/tree/main

## SQL - _____

You will use the `airlines` data from the SQL database that I used in the example in the [tutorial](https://advanced-ds-in-r.netlify.app/posts/2021-03-29-sqlinr/). Be sure to include the chunk to connect to the database here. And, when you are finished, disconnect. You may need to reconnect throughout as it times out after a while.

```{r}
con_air <- dbConnect(RMySQL::MySQL(), 
                     dbname = "airlines", 
                     host = "mdsr.cdc7tgkkqd0n.us-east-1.rds.amazonaws.com", 
                     user = "mdsr_public", 
                     password = "ImhsmflMDSwR")
```



**Tasks**:

1. Create a SQL chunk and an equivalent R code chunk that does the following: 

for each airport (with its name, not code), year, and month 
find the total number of departing flights, 
the distinct destinations to which they flew, 
the average length of the flight, 
the average distance of the flight, 
and the proportion of flights that arrived more than 20 minutes late. 

In the R code chunk, write this out to a dataset. (HINT: 1. start small! 2. you may want to do the R part first and use it to "cheat" into the SQL code).  

```{r}
dbListFields(con_air, "flights")

```


```{r}
overall_smry <-
  tbl(con_air, "flights") %>%
  group_by(origin, year, month) %>%
  summarize(num_flights = n(),
            num_destinations = n_distinct(dest),
            avg_flight_length = mean(air_time),
            avg_flight_distance = mean(distance),
            prop_late_over20 = mean(arr_delay > 20))

```


```{r}
show_query(overall_smry)
```





```{sql connection= con_air}
SELECT `origin`, `year`, `month`, 
  COUNT(*) AS `num_flights`, 
  COUNT(DISTINCT `dest`) AS `num_destinations`,
  AVG(`air_time`) AS `avg_flight_length`, 
  AVG(`distance`) AS `avg_flight_distance`, 
  AVG(`arr_delay` > 20.0) AS `prop_late_over20`
FROM `flights`
GROUP BY `origin`, `year`, `month`
```


  - With the dataset you wrote out, create a graph that helps illustrate the "worst" airports in terms of late arrivals. You have some freedom in how you define worst and you may want to consider some of the other variables you computed. Do some theming to make your graph look glamorous (those of you who weren't in my intro data science class this year may want to watch Will Chase's [Glamour of Graphics](https://www.youtube.com/watch?v=h5cTacaWE6I) talk for inspiration). 
  
```{r}
overall_smry_top10 <-  overall_smry %>%
  group_by(origin) %>%
  arrange(prop_late_over20, .by_group = TRUE) %>%
  top_n(10)
```

```{r}
overall_smry_top10
```

  
```{r}
overall_smry %>% 
  arrange(desc(prop_late_over20)) %>%
    top_n(10) %>%
  ggplot(aes(x = prop_late_over20,
             y = fct_reorder(origin, prop_late_over20, .desc=TRUE, ))) +
  geom_col(fill = "navy") +
  scale_x_continuous(expand = c(0,0), 
                     labels = scales::percent) +
  labs(x = NULL,
       y = NULL,
       title = "Which airport had the largest % of flights that \nwere more than 20 minutes late from 2010-2017?") +
  theme_minimal()
```
  
```{r}
overall_smry %>% 
  head(100) %>% 
  arrange(desc(prop_late_over20)) %>%
  top_n(10)
```

```{r}
flights_over20
```
  
  
  - Although your graph was truly inspirational, you've been requested to "boil it down to a few numbers." Some people just don't appreciate all that effort you put in. And, you need to use the already summarized data that you already pulled in from SQL. Create a table with 6 or fewer rows and 3 or fewer columns that summarizes which airport is the "worst" in terms of late arrivals. Be careful with your calculations. You may consider using the `kable`, `kableExtra`, or `gt` packages to make your table look truly spectacular.
  

```{sql connection= con_air}
SELECT `origin`, `year`, `month`, 
  COUNT(*) AS `num_flights`, 
  COUNT(DISTINCT `dest`) AS `num_destinations`,
  AVG(`air_time`) AS `avg_flight_length`, 
  AVG(`distance`) AS `avg_flight_distance`, 
  AVG(`arr_delay` > 20.0) AS `prop_late_over20`
FROM `flights`
GROUP BY `origin`, `year`, `month`
ORDER BY `prop_late_over20` DESC
LIMIT 3;
```

  
  
2. Come up with your own interesting question that data in the airlines database can help you answer. Write a SQL query and equivalent R code chunk to extract the data you need and create an elegant graph to help answer the question. Be sure to write down the question so it is clear. 


**Question: how does the avg delay performance of airlines compare over the years?** 

```{r}
my_data <-
  tbl(con_air, "flights") %>%
  group_by(carrier) %>%
  summarize(avg_delay = mean(arr_delay)) 

```


```{r}
show_query(my_data)
```


```{sql connection= con_air}
SELECT `carrier`, AVG(`arr_delay`) AS `avg_delay`
FROM `flights`
GROUP BY `carrier`
```


```{r}
my_data %>% 
  ggplot(aes(x = fct_reorder(avg_delay, median, .desc = TRUE),
             y = carrier, fill = navy)) +
  geom_col() +
  labs(x = NULL,
       y = NULL,
       title = "How does avg delay compare for carriers?") +
  theme_minimal()
```


```{r}
dbDisconnect(con_air)
```


## Shiny app - _______

If you are new to Shiny apps or it's been awhile since you've made one, visit the Shiny links on our course [Resource](https://advanced-ds-in-r.netlify.app/resources.html) page. 

Check out my Shiny tips [here](https://github.com/llendway/shiny_tips). 

Everyone should watch the [Theming Shiny](https://youtu.be/b9WWNO4P2nY) talk by Carson Sievert so you can make your app look amazing.

**Tasks:**

* Set up a separate project and GitHub repo for this app. The app needs to be created in a file called *exactly* app.R that is also in the project folder.  

* At the top of the file, load any libraries and data you will use in the app.  
* Use whatever data you'd like and create an app. It doesn't have to be super fancy, BUT it needs to incorporate all three of the new techniques I showed in the [Shiny tips](https://github.com/llendway/shiny_tips) - `reactive()`, referencing variable names, and updating a UI input with other UI input. 
* Use the `bslib` to theme your shiny app!  
* Publish your app to [shinyapps.io](https://www.shinyapps.io/). There are instructions for doing that on the tutorial I linked to above.   
* Write a paragraph or two describing your app on your website! Link to the app and your GitHub repository in your post. Include a link to your post here. 


## Function Friday problems -- _______


REMEMBER TO ADD YOUR GITHUB LINK AT THE TOP OF THE PAGE AND UNCOMMENT THE `knitr` OPTIONS.


